\documentclass[a4paper,10pt]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[tbtags]{amsmath}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{eucal}
\usepackage{hyperref, url}
\usepackage{float}
\usepackage{verbatim} % For verbatiminput.
\usepackage{fancyvrb}

\pdfinfo{            
          /Title      (T-61.3050 Machine Learning: Basic Principles)
          /Author     ()
          /Keywords   ()
}

\DefineVerbatimEnvironment {code}{Verbatim}
   {%numbers=left,numbersep=2mm,
     %frame=lines,framerule=0.1mm, 
     fontsize=\small}
\RecustomVerbatimCommand{\VerbatimInput}{VerbatimInput}
   {%numbers=left,numbersep=2mm,
     frame=lines,framerule=0.1mm, fontsize=\small}
\DefineVerbatimEnvironment {outputlog}{Verbatim}
  {frame=lines,framerule=0.1mm, fontsize=\small}

\newcommand{\XXX}[1]{{\bf XXX #1}}
\parindent 0mm
\parskip 3mm

%------------------------------------------------------------------------------
% First page
%------------------------------------------------------------------------------

% add your student number in parenthesis
\title{T-61.3050 Term Project, final report\\ % Dictated in instructions.
       Using decision trees for spam classification}
\author{Jori Bomanson (81819F) \\
  {\tt jori.bomanson@aalto.fi} \\
  \\
  Sami J. Lehtinen (44814P)\\ 
  {\tt sjl@iki.fi} \\
}
\begin{document}

\floatstyle{plain}
\newfloat{Listing}{t}{lol}
\floatname{Listing}{Listing}

\maketitle
\thispagestyle{empty}
\pagebreak
\pagenumbering{arabic}

\section{Abstract}
% Include the key points of your work in a few sentences (including methods
% used and conclusions).
We trained a univariate binary classification tree to detect spam email.


\section{Rationale}

The choice of using a decision tree to solve a task with boolean variables
seemed natural. Such a tree could definitely match whatever complexity there
would be in the training data.
In fact we figured our success would be determined mainly by how we would
manage to avoid overfitting.

Decision trees were also tempting in that they are much easier to
visualize than, e.g., naive bayes classifiers.  For example, see
\XXX{references to figures}

\section{Principles of decision trees}

Decision trees are a non-parametric method for data classification.
They don't try to approximate any model behind the data, instead relying
on tactic of ``similar inputs produce similar outputs''.

\subsection{Calculating impurity}

We chose to measure impurity with the entropy function \XXX{proper
citing} (Quinlan 1986)
\begin{equation*}
\begin{split}
\mathcal{I}_m &= - \sum_{i=1}^K p_m^i \log_2 p_m^i  \\
\text{for a node} \quad & m  \\
\text{where} \quad 0 \log 0 &\equiv 0  \\
p_m^i &= P(\text{Instance reaching node m belongs to class i})  \\
K &= 2
\end{split}
\end{equation*}

\subsection{Pruning the tree}

Decision trees can be trained to perfectly match the training
data\cite[\XXX{page number}]{alpaydin2004}.  This will lead the model to
gross overfitting, which manifests as a large generalization error.
This can be avoided by pruning the tree.  There are two non-exclusive
approaches to pruning, prepruning and postpruning.

Without pruning, a tree could look like \XXX{ref to non-pruned figure}
with a training set size of 1000.  

\XXX{figure of a tree without pruning}
\XXX{table of errors with a tree without pruning}

Classified as spam: 5556 / 9000
Validation:
 Classified as spam: 5556 / 9000
 Classified as spam in validation set: 5989 / 9000
 Correctly classified: 8499
 Correctly classified as spam: 5522
 Accuracy: 0.944
 Precision: 0.994
 Recall: 0.922
 Perplexity: 1.324

\subsubsection{Prepruning}

In prepruning, the recursion of the training algorithm is stopped once
certain limits are reached.  Examples include obtained impurity and number
of data rows.

In our first versions of the classifier we only did prepruning, which
resulted in very small and tidy trees, but the accuracy left comething
to be desired. \XXX{some data} Current model uses one of the suggested
approaches from \cite{alpaydin2004}, where we make the tree as pure as
possible (no prepruning) and handle overfitting by postpruning the tree.
Removing prepruning didn't have an adverse effect for the final
accuracy. \XXX{data contradicts, make a table.  Note training times for
  large datasets.}

Pre- and postpruned.

Classified as spam: 5744 / 9000
Validation:
 Classified as spam: 5744 / 9000
 Classified as spam in validation set: 5966 / 9000
 Correctly classified: 8594
 Correctly classified as spam: 5652
 Accuracy: 0.955
 Precision: 0.984
 Recall: 0.947
 Perplexity: 1.201

When prepruning, we stopped the algorithm by limiting by the results
from the entropy calculation, by columns (features) and rows (messages)
left in a recursion step.  For example, \XXX{more detailed explanation}.

\XXX{figure of a tree with large values in prepruning, prepruned-1000.dot}
\XXX{table of errors with a prepruned tree}

1000 items in training set, theta = 0.04, min\_row\_ratio = 0.002


Validation:
 Classified as spam: 5684 / 9000
 Classified as spam in validation set: 5962 / 9000
 Correctly classified: 8578
 Correctly classified as spam: 5612
 Accuracy: 0.953
 Precision: 0.987
 Recall: 0.941
 Perplexity: 1.268

\subsubsection{Postpruning}
\label{sect:postpruning}

Postpruning is used to reduce generalization error by using a a special
pruning data set to check whether removing a subtree from the decision
tree will result in the same or smaller error against the pruning set.

\XXX{figure of a postpruned tree}
\XXX{table of errors with a postpruned tree}

1000 items.
Chosen classifier has accuracy of 0.98.
Classified as spam: 5709 / 9000
Validation:
 Classified as spam: 5709 / 9000
 Classified as spam in validation set: 5952 / 9000
 Correctly classified: 8579
 Correctly classified as spam: 5620
 Accuracy: 0.953
 Precision: 0.984
 Recall: 0.944
 Perplexity: 1.182

6000 items, pre- and postpruning.
Validation:
 Classified as spam: 2604 / 4000
 Classified as spam in validation set: 2626 / 4000
 Correctly classified: 3890
 Correctly classified as spam: 2560
 Accuracy: 0.973
 Precision: 0.983
 Recall: 0.975
 Perplexity: 1.130

6000 items, just postpruning.
Validation:
 Classified as spam: 2601 / 4000
 Classified as spam in validation set: 2648 / 4000
 Correctly classified: 3907
 Correctly classified as spam: 2578
 Accuracy: 0.977
 Precision: 0.991
 Recall: 0.974
 Perplexity: 1.178

It should be noted that training the set without prepruning takes many
times longer than with it.

real	61m33.824s
user	57m22.444s
sys	1m15.947s

Prepruned tree takes a hit in accuracy, but is much faster to train.

Validation:
 Classified as spam: 2608 / 4000
 Classified as spam in validation set: 2683 / 4000
 Correctly classified: 3881
 Correctly classified as spam: 2586
 Accuracy: 0.970
 Precision: 0.992
 Recall: 0.964
 Perplexity: 1.233

real	14m28.909s
user	12m18.457s
sys	1m18.402s

\section{Validation of our approach}

\subsection{Accuracy, precision, recall, and perplexity}

The criteria for measuring the goodness of the classifier are the same
as the ones used in the data challenge.  As described in
\cite{termproject}, 

\subsection{K-fold cross validation}

We implemented 10-fold cross validation to our classifier.  This
gave more stable values for generalization error when compared to
repeatedly training the classifier without the cross validation.

One part of the training data set was used for postpruning the tree, see
section \ref{sect:postpruning}.

\subsection{Dummy model}


\section{Results with different training set sizes}

XXX results to a table.

1000 items.

\begin{outputlog}
Chosen classifier has accuracy of 1.0.
Classified as spam: 5777 / 9000
Validation:
 Classified as spam: 5777 / 9000
 Classified as spam in validation set: 6134 / 9000
 Correctly classified: 8475
 Correctly classified as spam: 5693
 Accuracy: 0.942
 Precision: 0.985
 Recall: 0.928

real	0m44.221s
user	0m44.040s
sys	0m0.150s
\end{outputlog}

4000 items in training set.

\begin{verbatim}
Chosen classifier has accuracy of 0.9725.
Classified as spam: 3984 / 6000
Validation:
 Classified as spam: 3984 / 6000
 Classified as spam in validation set: 4064 / 6000
 Correctly classified: 5798
 Correctly classified as spam: 3923
 Accuracy: 0.966
 Precision: 0.985
 Recall: 0.965

real	9m41.857s
user	9m40.900s
sys	0m0.520s
\end{verbatim}

6000 items in training set.

\begin{verbatim}
Chosen classifier has accuracy of 0.978333333333.
Classified as spam: 2677 / 4000
Validation:
 Classified as spam: 2677 / 4000
 Classified as spam in validation set: 2725 / 4000
 Correctly classified: 3886
 Correctly classified as spam: 2644
 Accuracy: 0.972
 Precision: 0.988
 Recall: 0.970

real	19m40.578s
user	19m37.980s
sys	0m1.100s
\end{verbatim}

\subsection{Training algorithm complexity}

\section{Conclusions}

The chosen approach didn't achieve the same high marks for accuracy as
the winners of the data challenge.  Tweaking of the approach is
naturally possible, but it seems that decision trees are more suited in
this as a part of combined classifier instead of on their own.  This
does not mean that the approach is completely without merits; decision
trees were applied here without prior knowledge to the distribution of
the spam messages.  In the data challenge, the winning teams had tweaked
the prior probabilities for spam to match the distribution in the data
set.  In the decision tree model, this is not not necessary, or even
possible.

Looking at the generated decision trees, it is easy to follow the
process by eye.  This, we think, helps a lot in validating the approach.

\section{Comments on project difficulty}

\begin{thebibliography}{9}
\bibitem{alpaydin2004}
  Ethem Alpaydin,
  \emph{Introduction to Machine Learning}.
  Massachusetts Institute of Technology, Cambridge, Massachusetts,
  1st edition,
  2004. 415 pages. ISBN 0-262-01211-1.
\bibitem{termproject}
  \emph{Term Project 2011: Spam or Ham?}.
  \href{https://noppa.aalto.fi/noppa/kurssi/t-61.3050/term\_project}
  {https://noppa.aalto.fi/noppa/kurssi/t-61.3050/term\_project}
  [Web article]. 2011-11-6. [Cited 2011-11-19].

%\bibitem{press07}
%  William H. Press, Saul A. Teukolsky, William T. Vetterling, Brian P. Flannery,
%  \emph{Numerical recipes - the art of scientific computing}.
%  Cambridge University Press, New York,
%  3rd edition,
%  2007. 1235 pages. ISBN 978-0-521-88068-8.
%\bibitem{mellin10a}
%  Ilkka Mellin, 
%  \emph{Todennäköisyyslaskenta ja tilastotiede: Kaavat}.
%  Otaniemi, 2010. 390 pages.
%\bibitem{mellin10b}
%  Ilkka Mellin, 
%  \emph{Tilastolliset taulukot}.
%  Otaniemi, 2010. 13 pages.
\end{thebibliography}

\appendix
\section{Python code for the classifier}
\verbatiminput{decisiontree.py}

\end{document}
